---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit this file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# biomasscan

<!-- badges: start -->
<!-- badges: end -->

The goal of `biomasscan` is to provide an easy, programmatic way of computing biomass from a tree list. The functions of this package were designed to mimic the functionality of Natural Resources Canada's online tool, available [here](https://apps-scf-cfs.rncan.gc.ca/calc/en/biomass-calculator). 

## Installation

You can install the development version of `biomasscan` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tesera/biomasscan")
```

## Example

Biomass can be estimated for a single tree, by passing the relevant attributes. For example, for a Jack pine, 15.2 meters tall and with a diameter of 21 centimeters at breast height, we get:

```{r example1}
library(biomasscan)

compute_biomass_single_tree(height = 15.2, diameter = 21, 
                            species = 'Jack pine')
```

As we can see, we get estimates of biomass (in kilograms) for four different components. We can turn this into a wide dataset by passing `wide = TRUE`:

```{r example2}
compute_biomass_single_tree(height = 15.2, diameter = 21, 
                            species = 'Jack pine', wide = TRUE)
```

We can also get the biomass estimates for a whole tree list:

```{r example3, message=FALSE}
library(tibble)

test_df <- tribble(
    ~spec, ~diam, ~ht,
    'Jack pine', 21, 15.2,
    'Trembling aspen', 32.1, 28,
    'Black spruce', 31, 22.6,
    'Black spruce', 21, 20.8,
    'Eucaplyptus', 14, 10.2
)

test_df

compute_biomass(test_df, species = 'spec', diameter = 'diam', 
                height = 'ht')
```


## References

  - Lambert, M.-C., C.-H. Ung, and F. Raulier (2005). *Canadian national biomass equations*. Can. J. For. Res 35: 1996-2018.
  - Ung, C.-H., Bernier, P., Guo, X.-J. (2008). *Canadian national biomass equations: new parameter estimates that include British Columbia data*. Can. J. For. Res 38:1123-2232.
